\documentclass[11pt]{gsasthesis} % 10,11 and 12pt fonts allowed

%%%%%%%%%%%%%%%% PACKAGES YOU PROBABLY WANT %%%%%%%%%%%%%%%%
% Include packages you want. The gsasthesis style file already includes
% packages "setspace" and "tocbibind".

\usepackage{etex} % extend the number of registers

% GSAS: "all margins should be at least 1 inch."
\usepackage[margin={1.2in}]{geometry}
% If you want asymmetric margins for two-sided documents, use the "twoside"
% option, as in
% \usepackage[top=1in,bottom=1.5in,left=1in,right=1.5in,twoside]{geometry} The
% left and right options become inner and outer margins The default horizontal
% latex margin ratio is 2:3. The default vertical top:bottom margin ratio is 2:3
% also. You can also set it directly by passing the hmarginratio option to the
% geometry package, as in
% \usepackage[top=1in,left=1in,vmarginratio=2:3,hmarginratio=2:5,twoside]{geometry}

% Appendix package. Not necessary, but it does make managing appendices easier
\usepackage[titletoc]{appendix}

%%%%%%%%%%%%%%%% PACKAGES MAY WANT %%%%%%%%%%%%%%%%

% sideways tables and figures
\usepackage{rotating}

% tables that spill over multiple pages
\usepackage{longtable}

% references
\usepackage{natbib}

% fonts that are nicer than defaults
\usepackage[sc]{mathpazo}
\usepackage{courier}

% Use 8-bit encoding that has 256 glyphs, pretty please
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% babel is required for blindtext, which generates random text
\usepackage[english]{babel}
\usepackage{blindtext}

% math support
\usepackage{amsmath}

% Slightly tweak font spacing for aesthetics
\usepackage{microtype}

\usepackage{subcaption}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage[plain]{algorithm}
\usepackage{algpascal}
\algdef{SE}{For}{End}[2]{%
  \textkeyword{for} \(#1\) \textkeyword{to} \(#2\) }{%
  \textkeyword{end}}
\algdef{SE}{While}{End}[1]{%
  \textkeyword{while} \(#1\)}{%
  \textkeyword{end}}
\algdef{SE}{If}{EndIf}[1]{%
  \textkeyword{if} \(#1\)}{%
  \textkeyword{end}}

\usepackage{amsfonts,color}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand\tab[1][1cm]{\hspace*{#1}}

\usepackage{array}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}

% You need the footmisc package with the stable option if you want to have
% footnotes inside section titles, for example to say that a particular chapter
% has been co-authored with someone. The multiple option ensures that there is a
% comma between two consecutive footnotes
\usepackage[stable,multiple]{footmisc}


% Nicer captions
\RequirePackage[font=small,format=plain,labelfont=bf,textfont=it]{caption}
\addtolength{\abovecaptionskip}{1ex}
\addtolength{\belowcaptionskip}{1ex}

%%%%%%%%%%%%%%%% COMPULSORY FIELDS %%%%%%%%%%%%%%%%

\title{Bayesian Optimization of a Wearable Assistive Device Using an Estimator Stopping Process} % needs to match title on DAC
\author{Charles Liu} % full name as it appears on your GSAS record, needs
                          % to match name on DAC
\degreename{Master of Engineering}
\degreefield{Computational Science \& Engineering} % Official name of subject as listed in GSAS
                                % handbook
\department{Institute of Applied Computational Science} % official name of department
\degreemonth{May} % Month of Defense (i.e. month when DAC was signed)
\degreeyear{2018} % Year the DAC was signed
\principaladvisor{Professor Scott Kuindersma}

% Optionally, you can add a second advisor, but you can't have three
% \secondadvisor{Professor George Secondary}



\begin{document}

%%%%%%%%%%%%%%%% FRONTMATTER %%%%%%%%%%%%%%%%

\pagenumbering{roman} % GSAS wants roman page numbers for frontmatter

% the following four pages are required in that order. The first two pages are
% not allowed to have page numbers, this is taken care of in the class file.
\thesistitlepage
\copyrightpage
\begin{abstract}
  Recent \emph{human-in-the-loop} (HIL) optimization studies using wearable devices have shown an improved average metabolic reduction by optimizing control parameters online during short-duration experiments. However, the coupling of slow metabolic dynamics, high measurement noise, and hard experimental time constraints creates significant practical challenges for scaling optimization methods to more expressive control strategies. Prior work on applying gradient descent and Bayesian optimization methods to perform HIL optimization have decoupled the estimation and parameter selection problems, which leads to fixed estimation intervals and imposes a hard limit on the number of parameter evaluations possible in a given time budget. In this work, a different approach is taken that couples estimation and parameter selection, allowing the algorithm to spend less time refining the metabolic estimates for parameters that are unlikely to improve performance over the best values observed so far. We represent this early stopping mechanism with two different models that are incorporated in a standard Bayesian optimization scheme using an unscented Kalman filter for estimating metabolic rate. Performance is analyzed in several numerical examples and in pilot experiments with two human subjects optimizing 6 free control parameters of a hip soft exosuit.
\end{abstract}

% Center headings for table of contents, LOT, and LOF and make them smaller so
% that "Abstract", "Acknowledgments" and "Contents" all look alike. Comment out
% if you want the default. If you want more control, use the "tocloft" package.
\renewcommand{\contentsname}{\protect\centering\protect\Large Contents}
\renewcommand{\listtablename}{\protect\centering\protect\Large List of Tables}
\renewcommand{\listfigurename}{\protect\centering\protect\Large List of Figures}
\renewcommand{\listalgorithmname}{\protect\centering\protect\Large List of Algorithms}

\tableofcontents % Table of contents

% The rest of the front matter: Lists of tables, figures, dedication and
% acknowledment is optional. Comment out whatever you don't like
\listoftables
\listoffigures
\listofalgorithms
\begin{acknowledgments}
  I am very grateful to Scott Kuindersma for being an incredible advisor. It has been a privilege to be his student and I appreciate his guidance, patience, and immense knowledge in helping me throughout the course of my study. It has also been a pleasure to work with Myunghee Kim, who was integral to the work in this thesis. Her wealth of experience in the field as well as overseeing the various levels of development was instrumental. I'd also like to thank Professor Conor Walsh, his lab and the Wyss staff, in particular Nikos Karavas, Sangjun Lee, Jinsoo Kim, Chih-Kang Chang, Asa Eckert-Erdheim, Maria Athanassiu, Brice Mikala Iwangou, Nicolas Menard, and Sarah Sullivan. This work would not have been possible without the hard work of everyone involved from the design and fabrication of the exosuit to the development of the controllers. Finally, I thank my parents, my friends, and the rest of my family for their amazing support.
\end{acknowledgments}


%%%%%%%%%%%%%%%% MAIN BODY %%%%%%%%%%%%%%%%
\pagenumbering{arabic} % reset page numbering and switch to arabic

% Introductory chapter. Comment out if you don't have an intro chapter, but I
% think most committees expect you to have one.
% Don't number the intro chapter, but add to to the table of contents
\addcontentsline{toc}{chapter}{Introduction}
\chapter*{Introduction}\label{ch:intro}
\input{intro}

\chapter{Bayesian Optimization}\label{ch:1}
\input{chapter1}

\chapter{Metabolic Cost Estimation}\label{ch:2}
\input{chapter2}

\chapter{Measurement Stopping Algorithms}\label{ch:3}
\input{chapter3}

\chapter{Results}\label{ch:4}
\input{chapter4}

\chapter*{Conclusion}\label{ch:concl}
\input{conclusion}

%%%%%%%%%%%%%%%% BACK MATTER %%%%%%%%%%%%%%%%

% Put appendices, bibliography, and supplemental materials here

% The bibliography may be single spaced within each entry, but must be
% double-spaced between each entry. Most bibliography styles leave space between
% entries, so that shouldn't be a problem.
\begin{singlespacing}
  % I like "References" better than "Bibliography"
  \renewcommand{\bibname}{References}

  % Any bibliohgraphy style that leaves space between entries is fine
  \bibliographystyle{ecca}
  \bibliography{references}
\end{singlespacing}

% Appendices from all chapters should go at the end
\input{appendix}

\end{document}
